% ------------------------------------------------------------
% ------------------------------------------------------------
  
% Insert the common KETCube AppNote Defines

% ------------------------------------------------------------
% ------------------------------------------------------------

\input{resources/appNotes/defs.tex}

\title{\UWBLogo KETCube AppNote 006:\\ KETCube Upstream Modules (\vhCurrentVersion)}

\author{Author: \vhListAllAuthorsLongWithAbbrev}
\date{Version \vhCurrentVersion\ from \vhCurrentDate}

% ------------------------------------------------------------
% ------------------------------------------------------------
  
% Insert the common KETCube AppNote Head

% ------------------------------------------------------------
% ------------------------------------------------------------

\input{resources/appNotes/head.tex}

% ------------------------------------------------------------
% ------------------------------------------------------------
  
% BEGIN of the KETCube appNote Content

% ------------------------------------------------------------
% ------------------------------------------------------------

\section*{About this Document}
\input{resources/about.tex}

This document describes KETCube upstream modules. 

\setcounter{tocdepth}{2}
\tableofcontents
\clearpage

\listoffigures
\listoftables
\begin{versionhistory}
% the star-marked versions should be merged into a single version when releasing KETCube
  \vhEntry{0.2.0*}{06.02.2021}{JB}{Integration of private modules and minor text update}
  \vhEntry{0.2.0*}{16.09.2019}{JB}{Initial version}
\end{versionhistory}
% history table ... do not number
\setcounter{table}{0}

\clearpage 
\pagenumbering{arabic} 
\pagestyle{headings} 

\clearpage

\section{Module Activation}\label{sec:activation}
KETCube is a modular platform, where number of modules is included in the platform release. Any module can be individually enabled or disabled based on user needs and extension boards connected to the KETCube {\it Main Board}. For details, please refer to the KETCube Datasheet \cite{ZCU:KETCube:05-2018}.

To display list of available modules, open the KETCube Terminal and use \docKCCmdInline{list} command. Commands \docKCCmdInline{enable}/\docKCCmdInline{disable} are used to turn ON/OFF modules (e.g. \docKCCmdInline{enable HDCX080}). When module is enabled, it starts to perform defined operation (e.g. measure RH and Temperature).

Debug messages are useful when device is initially configured or in case, that unexpected behavior occurred.

Debug messages are printed to serial terminal interface. They can be configured on the per-module basis by setting the {\it severity level} to the selected module or to the KETCube core.

Four severity levels are defined:
\begin{itemize}
  \item[0 --] NONE -- no messages enabled
  \item[1 --] ERROR -- only error messages enabled (default severity)
  \item[2 --] INFO -- error and informational messages enabled
  \item[3 --] DEBUG -- previous message groups and debug information are enabled
\end{itemize}

The second (optional) parameter of the “enable” command is used to configure the {\it severity level} of the selected module (e.g. \docKCCmdInline{enable HDCX080 3} enables debug messages for HDCX080 module). The command \docKCCmdInline{set core severity} is used to configure \docKCCmdInline{severity level} of the KETCube core and command \docKCCmdInline{set driver severity} is used to configure {\it severity level} of the KETCube low-level drivers.

\docNote{Debug messages may cause a situation, where it may be difficult to write commands (terminal echo is foggy due to lot of debug messages produced by KETCube)}.

The commands \docKCCmdInline{show(r)}/\docKCCmdInline{set(r)} are used to show/set KETCube (module) settings. The commands are organized in a tree-like manner: the \docKCCmdInline{show(r)}/\docKCCmdInline{set(r)} command is the root command, it is followed by the module name and by the module variable, which is to be set. 

\begin{docCodeExample}
\begin{verbatim}
>> show LoRa devEUI
[root command] [module name] [variable name]
\end{verbatim}
\end{docCodeExample}

\docNote{The persistent parameters are saved into on-chip EEPROM, the running configuration is stored in RAM.}

% ====================
\clearpage
\section{KETCube Actuation Modules}
\docNote{No actuation modules are currently part of the KETCube upstream.}
  
% ====================
\clearpage
\section{KETCube Communication Modules}

\subsection{LoRa Module}
\docKCModName{LoRa} allows to use KETCube as a LoRaWAN node device. In KETCube are supported two versions of the LoRaWAN protocol: 1.1.0 and 1.0.3. The LoRaWAN version is specified in compile-time and cannot be changed by user.

The \docKCModName{LoRa} can be enabled the same way as any other module -- see Section \ref{sec:activation}. Most of the LoRaWAN parameters can be set by using the KETCube Terminal interface.

Write command \docKCCmdInline{show LoRa} or \docKCCmdInline{set LoRa} to display commands related to the \docKCModName{LoRa}:

\begin{docCodeExampleTitled}{LoRa command set for LoRaWAN 1.0.3}
\begin{verbatim}
>> show LoRa
appEUI              LoRa application EUI
appKey              LoRa application key
appSKey             LoRa app session Key
devAddr             LoRa device address
devEUIBoard         Board (boardID-based) LoRa device EUI
devEUICustom        Custom LoRa device EUI
enableABP           Enable ABP
enableCustomDevEUI  Custom (user-defined) LoRa device EUI
nwkSKey             LoRa network session Key
txDatarate          Uplink datarate (0-15)

>> set LoRa
appEUI              LoRa application EUI
appKey              LoRa application key
appSKey             LoRa app session Key
devAddr             LoRa device address
devEUICustom        Custom LoRa device EUI
enableABP           Enable ABP
enableCustomDevEUI  Custom (user-defined) LoRa device EUI
nwkSKey             LoRa network session Key
txDatarate          Uplink datarate (0-15)
\end{verbatim}
\end{docCodeExampleTitled}

\begin{docCodeExampleTitled}{LoRa command set for LoRaWAN 1.1.0}
\begin{verbatim}
>> show LoRa
joinEUI             LoRa join EUI
appKey              LoRa application key
appSKey             LoRa app session Key
devAddr             LoRa device address
devEUIBoard         Board (boardID-based) LoRa device EUI
devEUICustom        Custom LoRa device EUI
enableABP           Enable ABP
enableCustomDevEUI  Custom (user-defined) LoRa device EUI
fNwkSIntKey         Forwarding Network session integrity key
nwkKey              Network key
nwkSEncKey          Network session encryption key
sNwkSIntKey         Serving Network session integrity key
txDatarate          Uplink datarate (0-15)

>> set LoRa
joinEUI             LoRa join EUI
appKey              LoRa application key
appSKey             LoRa app session Key
devAddr             LoRa device address
devEUICustom        Custom LoRa device EUI
enableABP           Enable ABP
enableCustomDevEUI  Custom (user-defined) LoRa device EUI
fNwkSIntKey         Forwarding Network session integrity key
nwkKey              Network key
nwkSEncKey          Network session encryption key
sNwkSIntKey         Serving Network session integrity key
txDatarate          Uplink datarate (0-15)
\end{verbatim}
\end{docCodeExampleTitled}

\subsubsection{LoRaWAN Node Activation}
To select or check OTAA/ABP mode settings, the following commands are used: \docKCCmdInline{set LoRa enableABP {0|1}}/\docKCCmdInline{show LoRa enableABP}. The “set” command parameter is boolean {\tt 0} or {\tt 1}. OTAA is the default mode.

\subsubsection{LoRaWAN devEUI}
User-defined devEUI (\docKCCmdInline{set LoRa enableCustomDevEUI 1} or manufacturer's devEUI (\docKCCmdInline{set LoRa enableCustomDevEUI 0}) can be used. The selected option can be checked via \docKCCmdInline{show LoRa enableCustomDevEUI} command. The user-defined devEUI can be configured by \docKCCmdInline{set LoRa devEUICustom} command. Manufacturer's devEUI is used by default.

\clearpage

\begin{docCodeExampleTitled}{Set user-defined devEUI}
\begin{verbatim}
>> set LoRa devEUICustom 1122334455667788
>> set LoRa enableCustomDevEUI 1
>> reload

>> show LoRa devEUICustom

Custom devEUI is displayed ...
\end{verbatim}
\end{docCodeExampleTitled}

\begin{docCodeExampleTitled}{Use Manufacturer's devEUI}
\begin{verbatim}
>> set LoRa enableCustomDevEUI 0
>> reload

>> show LoRa devEUIBoard

Board devEUI is displayed ...
\end{verbatim}
\end{docCodeExampleTitled}

\subsubsection{LoRaWAN OTAA Parameters}
The LoRaWAN appEUI and appKey can be set by using \docKCCmdInline{LoRa appEUI} and \docKCCmdInline{set LoRa appKey} commands and checked by using \docKCCmdInline{show LoRa appEUI} and \docKCCmdInline{show LoRa appKey} commands.
  
\begin{docCodeExampleTitled}{Set OTAA parameters}
\begin{verbatim}
>> set LoRa appEUI 1122334455667788
>> set LoRa appKey 11223344556677881122334455667788
\end{verbatim}
\end{docCodeExampleTitled}
  
\subsubsection{LoRaWAN ABP Parameters}
The LoRaWAN appSKey and nwkSKey can be set by using \docKCCmdInline{set LoRa appSKey} and \docKCCmdInline{set LoRa nwkSKey} commands and checked by using \docKCCmdInline{show LoRa appSKey} and \docKCCmdInline{show LoRa nwkSKey} commands. The static device address i configured via  \docKCCmdInline{set LoRa devAddr} command and checked via  \docKCCmdInline{show LoRa devAddr} command.
  
\begin{docCodeExampleTitled}{Set ABP parameters}
\begin{verbatim}
>> set LoRa devAddr DEADBEEF
>> set LoRa appSKey 1122334455667788
>> set LoRa nwkSKey 11223344556677881122334455667788
\end{verbatim}
\end{docCodeExampleTitled}

\subsubsection{LoRaWAN Period}
The period of data transmission is given by the {\it basePeriod} -- see the KETCube Datasheet \cite{ZCU:KETCube:05-2018}.

\subsubsection{LoRaWAN Tx Data Rate}
The Tx Data Rate is an important factor influencing the modulation, the spreading factor, and bandwidth. Lower numbers lead to a higher air-time, but higher transmission resilience, while higher numbers lead to a lower air-time, increased bitrate but lower transmission resilience.

\subsubsection{LoRaWAN Ports}

  The periodical messages are transmitted by using LoRaWAN port 2. Asynchronous messages (particular modules can indicate errors asynchronously, push-button event) are transmitted by using LoRaWAN port 3 if and only if the helper \docKCModName{AsyncTx} is enabled.
  
  If the helper module \docKCModName{RxDisplay} is enabled, the downlink messages received on the LoRaWAN port 10 are displayed in HEX format and those received on the LoRaWAN port 11 are displayed as a string (ASCII string expected).
  
  LoRaWAN port 12 is dedicated to custom ad-hoc LoRaWAN messaging implementations.
  
  LoRaWAN port 13 is dedicated for {\it KETCube Remote Terminal}.
  
  LoRaWAN port 14 is dedicated for the \docKCModName{UART2WAN} module.
  
  LoRaWAN port 14 is dedicated for the \docKCModName{Arduino} module.

% ---------------
\clearpage
\subsection{StarNetConcentrator Module}

\docWarn{This module is currently undocumented, as it remains in the alpha state!}

% ---------------
\clearpage
\subsection{StarNetNode Module}

\docWarn{This module is currently undocumented, as it remains in the alpha state!}

% ---------------
\clearpage
\subsection{TestRadio Module}

\docWarn{This module is intended for validation, not production! Please refer to source files, if you need to get a deep-in idea about its operation.}

\docNote{This module command set allows to invoke a continuous wave, set its power and frequency, or get the temperature of the radio module.}


% ====================
\clearpage
\section{KETCube Sensing Modules}

\subsection{ADC Module}
\docKCModName{ADC} allows to monitor DC Voltage at the KETCube AN PIN. It can be enabled the same way as any other module -- see Section \ref{sec:activation}.

\subsubsection{ADC Module Configuration}
  The \docKCModName{ADC} only allows to configure severity level, other configuration options are not available.
  
  The sample configuration procedure:
  
\begin{docCodeExample}
\begin{verbatim}
>> enable ADC
Executing command: enable

Setting module ADC:     success!
>> reload
Executing command: reload

Performing system reset and reloading KETCube configuration ...
\end{verbatim}
\end{docCodeExample}
  
  Once reloaded, the \docKCModName{ADC} measures connected voltage periodically. 
  
  For debugging purposes, it may be useful to watch measured voltage in terminal. The output for the severity level 2 (INFO) is shown below:
  
\begin{docCodeExample}
\begin{verbatim}
>> ADC :: Voltage@PA4: 133 mV
\end{verbatim}
\end{docCodeExample}

\subsubsection{ADC Module Output}
  The \docKCModName{ADC} output consist of one two-byte word (16-bit unsigned int):
  
\begin{tikzpicture}
 \node[draw,fill=SeaGreen3,minimum height=1cm,minimum width = 2cm,xshift=0cm](10/5){$ADC_{MSB}$};
 \node[draw,fill=SeaGreen2,minimum height=1cm,minimum width = 2cm,xshift=2cm](10/5){$ADC_{LSB}$};
\end{tikzpicture}

  Measured value represents the measured Voltage in millivolts. It is encoded in BigEndian format as unsigned integer.

% ---------------
\clearpage
\subsection{Arduino Module}

The Arduino module is the enabler of KETCube in Arduino IDE. We recently added the KETCube support to the Arduino IDE allowing us to create the KETCube module in Arduino IDE, upload and maintain the KETCube from the Arduino IDE environment.

By enabling this module, the Arduino-deployed code is activated.

\docWarn{This module and the Arduino IDE operation is currently documented in Czech only! See \url{http://edu.ketcube.cz/}}
  
% ------------------
\clearpage
\subsection{BatMeas Module}
  
The \docKCModName{BatMeas} module allows to monitor KETCube battery (if connected). 

The \docKCModName{BatMeas} can be enabled the same way as any other module -- see Section \ref{sec:activation}.

\subsubsection{BatMeas Module Configuration}
  Use the commands {\tt list} and {\tt bat} to select connected battery.
  
  The sample configuration procedure:
  
\begin{docCodeExample}
\begin{verbatim}
>> show batMeas list
Executing command: list
Available batteries:
0)       CR2032 (Up to 560mAh battery)
1)       LS33600 (15 Ah battery)
>> set batMeas bat 0
Executing command: bat

Write success!

>> enable BatMeas
Executing command: enable

Setting module BatMeas:     success!
>> reload
Executing command: reload

Performing system reset and reloading KETCube configuration ...
\end{verbatim}
\end{docCodeExample}

  Once reloaded, the \docKCModName{BatMeas} measures battery voltage periodically (voltage at KETCube 3V3 PIN). 
  
For debugging purposes, it may be useful to watch the \docKCModName{BatMeas} output in console. The output for the severity level 2 (INFO) is shown below:

\begin{docCodeExample}
\begin{verbatim}
>> batLevel :: 0
\end{verbatim}
\end{docCodeExample}

\subsubsection{BatMeas Module Output}
  The \docKCModName{BatMeas} output consist of a single byte only.
  
  The \docKCModName{BatMeas} output value (B) meaning is as follows:
  
  \begin{itemize}
    \item 0x00: battery is (almost) discharged
    \item 0x01 -- 0xFE: battery voltage scaled to 1 -- 254
    \item 0xFF: RFU
  \end{itemize}
  
  To compute actual battery voltage, use the following equation:
  
  $V = (V_{nom} - V_{min}) \cdot B/254 + V_{min} ~[mV]$
  
  $V_{min}$ is the lowest voltage value -- battery is considered as discharged. $V_{min}$ and $V_{nom}$ (nominal voltage) values for supported batteries are in the following table:
  
  \begin{table*}[!ht]
    %\hspace*{-4cm}
    \begin{tabular}{| p{3.5cm} | p{2cm} | p{2cm} |}
        \hline
        \rowcolor{SeaGreen3!30!} {\bf Battery} & $V_{min}$ [V] & $V_{nom}$ [V] \\
        \hline
        \hline
        CR2032 & 2.9 & 3.3 \\
        \hline
        LS33600 & 2.9 & 3.6 \\
        \hline
    \end{tabular}
    \addcontentsline{lot}{table}{Supported Batteries}
    \label{tab:spec:AMR}
   \end{table*}
  
% --------------

\clearpage
\subsection{BMEx80 Module}

The \docKCModName{BMEx80} module supports the BME280 and BME680 {\it Relative Humidity}, {\it Temperature} and {\it Preassure} sensors  allowing RH+T+P measurements.
  
\subsubsection{BMEx80 Module Configuration}
  
The \docKCModName{BMEx80} only allows to configure severity level, other configuration options are not available (some parameters can be configured in compile-time).
  
\begin{docCodeExample}
\begin{verbatim}
>> enable BMEx80 2

   ...
>> reload
Executing command: reload

Performing system reset and reloading KETCube configuration ...  
\end{verbatim}
\end{docCodeExample}
  
  Once reloaded, the \docKCModName{BMEx80} measures RH+T+P periodically. 
  
  For debugging purposes, it may be useful to watch RH+T output in console. The output for the severity level 2 (INFO) is shown below:

\begin{docCodeExample}
\begin{verbatim}
>> BMEx80 :: Type: BME 680, Temperature: 25.23, RH: 57.15, Pressure: 1112.35
\end{verbatim}
\end{docCodeExample}

\subsubsection{BMEx80 Module Output}
  The \docKCModName{BMEx80} output consist of 4 bytes:
  
\begin{tikzpicture}
 \node[draw,fill=SeaGreen3,minimum height=1cm,minimum width = 2cm,xshift=0cm](10/5){$RH_m$};
 \node[draw,fill=white,minimum height=1cm,minimum width = 2cm,xshift=2cm](10/5){$T_m$};
 \node[draw,fill=SkyBlue,minimum height=1cm,minimum width = 2cm,xshift=4cm](10/5){$P_{MSB}$};
 \node[draw,fill=SkyBlue,minimum height=1cm,minimum width = 2cm,xshift=6cm](10/5){$P_{LSB}$};
\end{tikzpicture}

  Relative Humidity is encoded as unsigned integer. The value units are \% RH.
  
  To compute actual RH, use the following equation:
  
  $RH = RH_m / 2 ~[\%]$

  Temperature is encoded as unsigned integer in additive code. The value units are degrees of Celsius.
  
  To compute actual temperature, use the following equation:
  
  $T = T_m / 2 - 40 ~[°C]$
  
  Preassure is encoded as unsigned integer. The value units are hecto Pascals.
  
  To compute actual preassure, use the following equation:
  
  $P = (P_{MSB} << 8 ~|~ P_{LSB}) / 2 ~[hPa]$
  
  
  An error for temperature, humidity and preassure is indicated by out-of-the range temperature/humidity/preassure values, particularly 0xFF or 0xFFFF values are used.
  
% --------------

\clearpage
\subsection{HDCX080 Module}

The HDC1080 and HDC2080 are {\it Relative Humidity} and {\it Temperature} sensors allowing RH+T measurement. One of them is assembled on-board. The HDCX080 module can be enabled the same way as any other module -- see Section \ref{sec:activation}.

\subsubsection{HDCX080 Module Configuration}
  The \docKCModName{HDCX080} only allows to configure severity level, other configuration options are not available.
  
\begin{docCodeExample}
\begin{verbatim}
>> enable HDCX080 2

   ...
>> reload
Executing command: reload

Performing system reset and reloading KETCube configuration ...  
\end{verbatim}
\end{docCodeExample}
  
  Once reloaded, the \docKCModName{HDCX080} measures RH+T periodically. 
  
  For debugging purposes, it may be useful to watch RH+T output in console. The output for the severity level 2 (INFO) is shown below:

\begin{docCodeExample}
\begin{verbatim}
>> HDCX080 :: Temperature: 25, RH: 57
\end{verbatim}
\end{docCodeExample}

\subsubsection{HDCX080 Module Output}
  The \docKCModName{HDCX080} output consist of 4 bytes:
  
\begin{tikzpicture}
 \node[draw,fill=SeaGreen3,minimum height=1cm,minimum width = 2cm,xshift=0cm](10/5){$T_{MSB}$};
 \node[draw,fill=SeaGreen3,minimum height=1cm,minimum width = 2cm,xshift=2cm](10/5){$T_{LSB}$};
 \node[draw,fill=white,minimum height=1cm,minimum width = 2cm,xshift=4cm](10/5){$RH_{MSB}$};
 \node[draw,fill=white,minimum height=1cm,minimum width = 2cm,xshift=6cm](10/5){$RH_{LSB}$};
\end{tikzpicture}

  Temperature is encoded in BigEndian as unsigned integer in additive code. The value units are degrees of Celsius.
  
  To compute actual temperature, use the following equation:
  
  $T = ((T_{MSB} << 8 ~|~ T_{LSB}) - 10000) / 10 ~[°C]$
  
  Relative Humidity is encoded in BigEndian format as unsigned integer in additive code. The value units are \% RH.
  
  To compute actual RH, use the following equation:
  
  $RH = (RH_{MSB} << 8 ~|~ RH_{LSB}) / 10 ~[\%]$
  
  An error for both temperature and humidity is indicated by out-of-the range temperature/humidity values, particularly 0xFFXX values are used, where XX indicates the error source:

\begin{itemize}
  \item {\tt 0x01} -- sensor (auto)detection failed
  \item {\tt 011} -- I2C write error
  \item {\tt 012} -- I2C write address error
  \item {\tt 021} -- I2C read error
\end{itemize}
  
% --------------

\clearpage
\subsection{ICS43432 Module}

\docWarn{This module is orphaned! It will be removed from the upstream release in a near future!}

% --------------

\clearpage
\subsection{LIS2HH12 Module}

\docWarn{This module is orphaned! It will be removed from the upstream release in a near future!}

% ---------------
\clearpage
\subsection{UART2WAN Module}

\docWarn{This module is currently undocumented, as it remains in the alpha state!}

% ------------------------------------------------------------
% ------------------------------------------------------------
  
% END of the KETCube appNote Content

% ------------------------------------------------------------
% ------------------------------------------------------------



% ------------------------------------------------------------
% ------------------------------------------------------------
  
% Insert the common KETCube appNote Tail

% ------------------------------------------------------------
% ------------------------------------------------------------

\input{resources/appNotes/tail.tex}

